---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

suppressPackageStartupMessages(library(lc, quietly = TRUE))
suppressPackageStartupMessages(library(magrittr, quietly = TRUE))
suppressPackageStartupMessages(library(pmatch, quietly = TRUE))
```

# lc -- List comprehension in R

[![Travis build status](https://travis-ci.org/mailund/lc.svg?branch=master)](https://travis-ci.org/mailund/lc)
[![AppVeyor build status](https://ci.appveyor.com/api/projects/status/wsopc251n1jpj40j/branch/master?svg=true)](https://ci.appveyor.com/project/mailund/lc/branch/master)
[![Coverage Status](https://img.shields.io/codecov/c/github/mailund/lc/master.svg)](https://codecov.io/github/mailund/lc?branch=master)
[![Coverage Status](https://coveralls.io/repos/github/mailund/lc/badge.svg?branch=master)](https://coveralls.io/github/mailund/lc?branch=master)
[![lifecycle](https://img.shields.io/badge/lifecycle-maturing-blue.svg)](https://www.tidyverse.org/lifecycle/#maturing)

The goal of `lc` is to implement Haskell- and Python-like list comprehension in R. List comprehensions provide a syntax for mapping and filtering sequences. In R we would use functions such as `Map` or `Filter`, or the `purrr` alternatives, for this, but in languages such as Haskell or Python, there is syntactic sugar to make combinations of mapping and filtering easier to program.


## Installation

You can install lc from GitHub with:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("mailund/lc")
```

or from CRAN using

```{r cran-installation, eval = FALSE}
install.packages("lc")
```


## Examples

### Quick-sort

The algorithm [quick-sort](https://en.wikipedia.org/wiki/Quicksort) is based on the idea that to sort a list you pick a random element in it, called the *pivot*, split the data into those elements smaller than the pivot, equal to the pivot and larger than the pivot. Then sort those smaller and larger elements recursively and concatenate the three lists to get the final sorted list. One way to implement this in R is using the `Filter` function:

```{r}
qsort <- function(lst) {
  n <- length(lst)
  if (n < 2) return(lst)

  pivot <- lst[[sample(n, size = 1)]]
  smaller <- Filter(function(x) x < pivot, lst)
  equal <- Filter(function(x) x == pivot, lst)
  larger <- Filter(function(x) x > pivot, lst)
  c(qsort(smaller), equal, qsort(larger))
}

(lst <- sample(1:10))
unlist(qsort(lst))
```

Using `lc`, you can implement the same algorithm like this:

```{r}
qsort <- function(lst) {
  n <- length(lst)
  if (n < 2) return(lst)
  
  pivot <- lst[[sample(n, size = 1)]]
  smaller <- lc(x, x = lst, x < pivot)
  equal <- lc(x, x = lst, x == pivot)
  larger <- lc(x, x = lst, x > pivot)
  
  c(qsort(smaller), equal, qsort(larger))
}

(lst <- sample(1:10))
unlist(qsort(lst))
```

### Eratosthenes sieve

The [Eratosthenes sieve](https://en.wikipedia.org/wiki/Sieve_of_Eratosthenes) algorithm identifies the primes in the first `n` natural numbers by removing those elements that are divisible by smaller numbers. We can implement it using `lc` like this:

```{r}
get_primes <- function(n) {
  not_primes <- lc(seq(from = 2*x, to = n, by = x), x = 2:sqrt(n)) %>% unlist %>% unique
  lc(p, p = 1:n, !(p %in% not_primes)) %>% unlist
}
get_primes(100)
```

Here, we create a list of all the non-primes first and then identifies the primes as those that are left. Alternatively, we can remove the non-primes iteratively while identifying the primes. For the implementation of this idea, we need to grow a list of primes. We don't want to do that using `list`, since this will give us a quadratic time algorithm, so I use the [`pmatch` package](https://github.com/mailund/pmatch) to implement a linked list first:

```{r}
library(pmatch)

linked_list := NIL | CONS(car, cdr : linked_list)

ll_length <- function(x, acc = 0) {
  cases(x,
        NIL -> acc,
        CONS(car,cdr) -> ll_length(cdr, acc + 1))
}
ll_to_list <- function(x) {
  y <- vector("list", length = ll_length(x))
  f <- function(x, i) {
    cases(x,
          NIL -> NULL,
          CONS(car, cdr) -> {
            y[[i]] <<- car
            f(cdr, i + 1)
        })
  }
  f(x, 1)
  y
}
```

The `get_primes` function can now be implemented like this:

```{r}
get_primes <- function(n) {
  candidates <- 2:n
  primes <- NIL
  while (length(candidates) > 0) {
    p <- candidates[[1]]
    primes <- CONS(p, primes)
    candidates <- lc(x, x = candidates, x %% p != 0)
  }
  primes %>% ll_to_list %>% unlist %>% rev
}
get_primes(100) 
```
